% M5 ablation subsection — DRAFT (fill 5-seed numbers after run_m5_full_ablation.py completes)
% Placement: after Section "Semantic Safety-Layer Weight Sweep" or as a new subsection in Results.
% Core message: this ablation confirms the paper's central claim that the SAC actor
% reads LLM-derived semantic features from the state, rather than relying solely on
% reward shaping or input-dimension expansion.

\subsection{Semantic-Source Ablation: Does the Actor Read LLM Semantics?}
\label{sec:m5_ablation}

A reviewer-relevant question is whether the gain of LE-DRL-SAC over SAC-Numeric is
attributable to the actor genuinely reading the LLM-derived semantic features from
the augmented state, or to indirect effects such as a larger input vector or a denser
training reward. We isolate these mechanisms with three diagnostic baselines and
three semantic sources under the main configuration ($w=0.9$, $\lambda_{\mathrm{sem}}=3.0$,
S1--S4, five seeds).

\textbf{Three baselines.} (i) \emph{LE-DRL-SAC} (native): the actor reads the real
semantic vector $s^{\mathrm{sem}}_t$ in $s^{\mathrm{aug}}_t$, and the semantic-consistency
regularizer (Eq.~6) actively pulls the actor toward a semantic prior action.
(ii) \emph{LE-DRL w/o Text}: the same 12-dim input is kept, but the five semantic
dims are clamped to zero; the regularizer still fires but its target degenerates to a
text-free prior. This isolates the value of the semantic \emph{content} from the value
of the larger input vector. (iii) \emph{SAC-Numeric}: a 7-dim input with no semantic
channel at all, isolating the value of the channel itself.

\textbf{Three semantic sources.} The $s^{\mathrm{sem}}_t$ fed to (i) and (ii) is drawn
from one of: (a) DeepSeek (the main-paper source), (b) a local keyword encoder
(\texttt{LocalSemanticEncoder}, a deterministic Chinese-keyword scorer), or (c) the
DeepSeek scores perturbed by clipped Gaussian noise ($\sigma{=}0.20$ on risk dims,
$\sigma{=}0.30$ on storage bias). Source (b) tests whether the gain is unique to LLM
understanding or matched by structured text features; source (c) tests whether the
semantic \emph{content} rather than the dimensionality carries the value.

\textbf{Result 1 — the actor reads the semantic state.} Table~\ref{tab:m5_ablation}
shows that under the main configuration LE-DRL-SAC (native) outperforms LE-DRL w/o
Text on every known-event scenario under the DeepSeek source: S1 $+491.7$ yuan
(95\% CI $[+433.5, +535.2]$, 5/5 seeds), S2 $+492.5$ ($[+480.0, +505.4]$, 5/5),
S3 $+960.6$ ($[+910.4, +997.3]$, 5/5), and S4 $+1{,}224.7$ ($[+1{,}158.5, +1{,}261.0]$,
5/5). Pooled across the 20 matched (seed, scenario) pairs the gap is $+792.4$ yuan
(95\% CI $[+654.7, +934.0]$, 20/20 favorable). Because the two baselines share the
same 12-dim input and the same regularizer activation, the only difference is whether
the semantic dims carry real scores or are zeroed. The consistent, all-seed-favorable
advantage of native therefore cannot be attributed to input-dimension expansion and
confirms that the SAC actor, under the semantic-consistency regularizer, learns to
read the LLM-derived features in $s^{\mathrm{sem}}_t$ and act on them. This is the
direct evidence for the language-enhanced mechanism named in the title. The advantage
is scenario-dependent and largest exactly where textual semantics should matter
most---S4 renewable-curtailment ($+1{,}225$) and S3 price-spike ($+961$)---consistent
with the main-paper observation in Section~\ref{sec:results}.

\textbf{Result 2 — the semantic content matters, not just the dimensionality.}
The noisy-source variant perturbs the DeepSeek scores with clipped Gaussian noise
while keeping the same five dims active. Under noise, LE-DRL-SAC reward falls
\emph{below} the zeroed-semantics baseline: the pooled native-minus-w/o-Text gap
reverses to $-1{,}429.1$ yuan (95\% CI $[-2{,}170.5, -548.5]$, 5/20 favorable).
In other words, semantic dims carrying \emph{wrong} values are worse than semantic
dims carrying \emph{no} values: the actor, having been trained by the
semantic-consistency regularizer to trust the semantic channel, is misled when
that channel is corrupted. This confirms that the actor exploits the semantic
\emph{values}, not merely the presence of the channel. The DeepSeek-minus-noisy
pooled gap is $+2{,}221.5$ yuan (95\% CI $[+1{,}382.7, +2{,}945.0]$, 15/20),
reinforcing that genuine semantic content is load-bearing.

\textbf{Result 3 — DeepSeek vs.\ keyword encoder on known events.} The local
keyword encoder (\texttt{LocalSemanticEncoder}) is a deterministic Chinese-keyword
scorer hand-designed for the four known event categories. Under the main
configuration it performs comparably to DeepSeek: the pooled native gap
(DeepSeek minus keyword) is $-28.4$ yuan (95\% CI $[-63.6, +8.3]$, 6/20 favorable),
statistically indistinguishable from zero. This indicates that for the four
\emph{known} event categories, which the keyword encoder was designed to cover,
structured text-derived scores are sufficient and the LLM offers no distinct
advantage. The LLM's distinct value is exposed precisely outside that coverage:
on the negative-price event S5 (Section~\ref{sec:unseen_event}), the keyword
encoder returns renewable-curtailment $0.0$ and storage-bias $-0.75$ (it has no
keyword for negative prices), whereas DeepSeek returns $0.9$ and $+0.8$---the
LLM recognizes a surplus event the keyword encoder mis-scores. The S5
new-event adaptation is therefore where language understanding, rather than
keyword matching, becomes load-bearing.

\textbf{Scope.} This ablation is run on the four known-event scenarios under the
main configuration ($\lambda_{\mathrm{sem}}=3.0$, $w=0.9$). The new-event adaptation
(S5) uses a relaxed regularizer ($\lambda_{\mathrm{sem}}=0$) that intentionally
frees the actor from the semantic target so it can learn cross-interval SOC
management; under that relaxed regime the native-vs-w/o-Text gap shrinks because
the regularizer that makes the actor read semantics is turned off---this is the
intended trade-off documented in Section~\ref{sec:unseen_event}, not a
contradiction of Result~1.

\textbf{Result 4 --- DeepSeek distinctness on a physics-reversal event (S8).}
Results 1--3 showed that on S1--S4 the DeepSeek and keyword sources are
statistically indistinguishable, and on S5 (negative-price) and S7
(export-curtailed local absorption) the physical signal alone is strong enough
to drive the correct action, washing out the semantic marginal contribution.
A reviewer-relevant question is whether there exists any event where the LLM's
semantic understanding is \emph{necessary}---where the correct response cannot be
learned from the physical signal alone. We construct such an event, S8
(frequency-reserve request), where (i) the correct response (keep SOC mid-band
for bidirectional reserve) is \emph{opposite} to what the physical signal
suggests (PV surplus $\to$ charge $\to$ fill up), (ii) reserve-request intervals
are scattered across all hours so the hour-of-day feature alone cannot predict
them, and (iii) a reserve-capacity penalty $\rho(\mathrm{SOC}-0.5)^2$ is applied
\emph{only} during intervals whose event text indicates a reserve request, making
``keep mid-band'' reward-contingent on the textual event. Under the main
configuration, LE-DRL-SAC with DeepSeek semantic features outperforms the keyword
encoder by $+788.6$ yuan (95\% CI $[+743.3, +837.5]$, 5/5 seeds favorable), and
pays a lower reserve penalty ($1{,}803.7$ vs.\ $2{,}337.5$ yuan). The keyword
encoder returns all-zero scores for the reserve event (its keyword table has no
``frequency/reserve'' branch), so it cannot steer the actor away from filling up
during reserve intervals; DeepSeek recognizes the reserve request and returns
renewable-curtailment $0.20$ and storage-bias $0.0$ (no charge bias, keep
mid-band). SAC-Numeric, which has no event-text channel at all, pays the highest
reserve penalty ($3{,}746.4$) and keeps mid-band only $2.5\%$ of reserve steps.
This is the scenario where the language-enhanced mechanism named in the title
becomes load-bearing: when the correct response reverses the physical signal and
the event timing is unpredictable from numeric features alone, the LLM's
semantic understanding yields a statistically significant advantage that the
keyword encoder cannot match.

\begin{table}[t!]
\caption{S8 Frequency-Reserve Distinctness, Main Config ($w=0.9$, $\lambda_{\mathrm{sem}}=3.0$, $\rho=600$), Five-Seed Mean}
\label{tab:m5_s8}
\centering
\begin{tabular}{llrrr}
\toprule
\textbf{Source} & \textbf{Model} & \textbf{Reward} & \textbf{Reserve cost} & \textbf{Mid-band rate} \\
\midrule
DeepSeek & LE-DRL-SAC & \textbf{-203,876.1} & \textbf{1,803.7} & 0.225 \\
Keyword & LE-DRL-SAC & -204,664.7 & 2,337.5 & 0.225 \\
DeepSeek & SAC-Numeric & -207,512.1 & 3,746.4 & 0.025 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t!]
\caption{Semantic-Source Ablation on S1--S4, Main Config ($w=0.9$, $\lambda_{\mathrm{sem}}=3.0$), Five-Seed Mean Reward (yuan)}
\label{tab:m5_ablation}
\centering
\begin{tabular}{llrrrr}
\toprule
\textbf{Source} & \textbf{Model} & \textbf{S1} & \textbf{S2} & \textbf{S3} & \textbf{S4} \\
\midrule
\multirow{3}{*}{DeepSeek} & LE-DRL-SAC (native) & -88,561.1 & -94,903.9 & -95,553.3 & -85,422.1 \\
 & LE-DRL w/o Text & -89,052.7 & -95,396.4 & -96,513.9 & -86,646.8 \\
 & SAC-Numeric & -89,052.9 & -95,397.0 & -96,514.3 & -86,647.2 \\
\midrule
\multirow{3}{*}{Keyword} & LE-DRL-SAC (native) & -88,532.3 & -94,787.9 & -95,565.9 & -85,440.6 \\
 & LE-DRL w/o Text & -89,052.7 & -95,396.4 & -96,513.9 & -86,646.8 \\
 & SAC-Numeric & -89,052.9 & -95,397.0 & -96,514.3 & -86,647.2 \\
\midrule
\multirow{3}{*}{Noisy} & LE-DRL-SAC (native) & -91,552.6 & -98,157.9 & -94,666.3 & -88,949.3 \\
 & LE-DRL w/o Text & -89,052.7 & -95,396.4 & -96,513.9 & -86,646.8 \\
 & SAC-Numeric & -89,052.9 & -95,397.0 & -96,514.3 & -86,647.2 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[t!]
\caption{Native (LE-DRL-SAC) minus w/o-Text gap, pooled over 20 (seed, scenario) pairs, 95\% paired-bootstrap CI}
\label{tab:m5_gaps}
\centering
\begin{tabular}{lrrr}
\toprule
\textbf{Semantic source} & \textbf{Gap (yuan)} & \textbf{95\% CI} & \textbf{Favorable} \\
\midrule
DeepSeek & \textbf{+792.4} & $[+654.7, +934.0]$ & 20/20 \\
Keyword & \textbf{+820.8} & $[+700.2, +946.7]$ & 20/20 \\
Noisy ($\sigma{=}0.20$) & -1,429.1 & $[-2{,}170.5, -548.5]$ & 5/20 \\
\bottomrule
\end{tabular}
\end{table}

% Bootstrap CI note: use the seed-paired paired-bootstrap (20,000 resamples) over
% the 5x4=20 matched (seed, scenario) pairs, consistent with the main-paper CI
% method. Report native-minus-w/o-Text and DeepSeek-minus-keyword gaps.
